---
title: 'N741: Homework 5'
author: "Craig Alder"
date: "March 15, 2017"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

## Load dataset in from `car` package

```{r}
library(car)
data(Wong)

library(dplyr)
# add an age group variable
Wong$agegrp <- case_when(
  (Wong$age > 0 & Wong$age <= 10) ~ 1,
  (Wong$age > 10 & Wong$age <= 20) ~ 2,
  (Wong$age > 20 & Wong$age <= 30) ~ 3,
  (Wong$age > 30 & Wong$age <= 40) ~ 4,
  (Wong$age > 40 & Wong$age <= 50) ~ 5,
  (Wong$age > 50 & Wong$age <= 60) ~ 6,
  (Wong$age > 60 & Wong$age <= 70) ~ 7,
  (Wong$age > 70 & Wong$age <= 100) ~ 8)

# convert to factor, add code levels and labels
Wong$agegrp <- factor(Wong$agegrp,
                  levels = c(1,2,3,4,5,6,7,8),
                  labels = c("Ages 1-10",
                             "Ages 11-10",
                             "Ages 21-10",
                             "Ages 31-10",
                             "Ages 41-10",
                             "Ages 51-10",
                             "Ages 61-70",
                             "Ages 71-100"))
```

1. Make a table of non-parametric statistics (median and IQR) for the number of days and duration grouped by `sex`. You'll be using `summarise()` from the `dplyr` package. For a given variable `x` you'll use `median(x, na.rm=TRUE)`, `quantile(x, 0.25, na.rm=TRUE)`, and `quantile(x, 0.75, na.rm=TRUE)`. Give the table a title using the `caption=` option and update the column names with something nice using the `col.names=` option in the `knitr::kable()` command. 

```{r}
tbl1 <- Wong %>%
group_by(sex) %>%
  summarise(Q1days = quantile(days, 0.25, na.rm=TRUE),
            mediandays = median(days, na.rm=TRUE),
            Q3days = quantile(days, 0.75, na.rm=TRUE),
            Q1duration = quantile(duration, 0.25, na.rm=TRUE),
            medianduration = median(duration, na.rm=TRUE),
            Q3duration = quantile(duration, 0.75, na.rm=TRUE))
  
knitr::kable(tbl1,
             col.names=c("Sex",
                         ".25Q Days","Median Days",".75Q Days",
                         ".25Q Duration","Median Duration",
                         ".75Q Duration"),
             caption="Days and Duration by Sex")

```

2.  Make a table of parametric statistics (mean and SD) for the performance outcomes `piq` and `viq` grouped by `sex`. Like the table above, you'll be using `summarise()` from the `dplyr` package. Now you'll use `mean(x, na.rm=TRUE)` and `sd(x, na.rm=TRUE)`. Give the table a title using the `caption=` option and update the column names with something nice using the `col.names=` option in the `knitr::kable()` command. 

```{r}
tbl2 <- Wong %>%
group_by(sex) %>%
  summarise(piqmean = mean(piq, na.rm=TRUE),
            piqsd   = sd(piq, na.rm=TRUE),
            viqmean   = mean(viq,na.rm=TRUE),
            viqsd  = sd(viq, na.rm=TRUE))
  
knitr::kable(tbl2,
             col.names=c("Sex",
                         "Math IQ, Mean","Math IQ SD",
                         "Verbal IQ, Mean","Verbal IQ,SD G2"),
             caption="IQ Scores by Sex")
```

3. Make a table containing the frequencies and relative percentages for `agegrp`. Use the example we did in class to help guide you.

```{r}
# find the sample size or number of rows in dataset
ss <- length(Wong$agegrp)

# group by each level of Medu
# find the raw counts using n()
# then compute the relative % by dividing by ss
tbl3 <- Wong %>%
  group_by(agegrp) %>%
  summarise(freq = n(),
            pct = n()*100/ss)

# make a summary table
knitr::kable(tbl3,
             col.names = c("Age Group",
                           "Frequency",
                           "Percent"),
             caption="Frequency Table for Age Group")
```

4. Make a regression model (Model 1) for the performance IQ (`piq`) using `age` and `sex`. Put the regression model results into a table.

```{r}
m1 <- lm(piq~age+sex, data=Wong)
sm1 <- summary(m1)
knitr::kable(sm1$coefficients,
             caption="Model 1: Math IQ regressed on age and sex")
```

5. Make a second regression model (Model 2) for performance IQ (`piq`) using `age` and `sex` plus `days` and `duration`. Put the regression model results into a table.

```{r}
m2 <- lm(piq~age+sex+days+duration, data=Wong)
sm2 <- summary(m1)
knitr::kable(sm2$coefficients,
             caption="Model 1: Math IQ regressed on age, sex, days, duration")
```

6. Finally, make a table showing the results from the `anova()` command comparing Model 1 and Model 2 you made above using the example we did in class as a guide. 

```{r}
anova(m1,m2)
```

7. STUDENT CHOICE - pick either a `htmlwidget` from [http://gallery.htmlwidgets.org/](http://gallery.htmlwidgets.org/) or do a "flexdashboard" using the templates at [http://rmarkdown.rstudio.com/flexdashboard/](http://rmarkdown.rstudio.com/flexdashboard/) as a guide.

```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
# Build the plot
mathIQ_plot <- ggplot(data = Wong, aes(x="Verbal IQ", y="Math IQ")) + 
  geom_point(aes(color=sex), size=2)
mathIQ_plot
ggplotly(mathIQ_plot)
```

Code I tried

{r setup, include=FALSE}
library(ggplot2)
library(plotly)
# Build the plot
mathIQ_plot <- ggplot(data = Wong, aes(x="Verbal IQ", y="Math IQ")) + 
  geom_point(aes(color=sex), size=2)
mathIQ_plot
ggplotly(mathIQ_plot)


### References

Wong, P. P., Monette, G., and Weiner, N. I. (2001) Mathematical models of cognitive recovery. Brain Injury, 15, 519–530.

Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.


The link to my Github account is [https://github.com/craigalder](https://github.com/craigalder). The link to my repository for this assignment is [https://github.com/craigalder/N741interactive](https://github.com/craigalder/N741interactive).